---
title: "Rens_HW_Wk7"
output: html_document
date: "2024-03-02"
---

## Load Libraries

```{r Loading Libraries}
# Load libraries
library(rio)
library(dplyr)
library(ggplot2)
library(factoextra)
library(cluster)
```

## Import and View the Data

```{r Importing Data}
# Import and view the data
data <- import("~/School/Predictive Modeling/CreditCards.csv")

head(data)
summary(data)
dim(data)
```

Looks like a lot of numbers to me. The instructions say to remove the CUST_ID column, but I don't see any column by that name. Judging by the information on the Kaggle web page I am going to make an educated guess that the "V1" column is a re-naming of the CUST_ID column. It's just a sequentially increasing variable. I'll go ahead and remove the "V1" column.

```{r Removing V1}
# Removed the "V1" column
data <- subset(data, select = -c(V1))
summary(data)
```

Looks good!

## Clustering

Let's determine the ideal number of clusters for the model.

```{r Finding Optimal Cluster Count}
# Model the appropriate number of clusters
fviz_nbclust(data, kmeans, method="gap_stat")
```

...aaannnnnd we're back in action. That took a little while. This is telling me that the optimal number of clusters is 2, so lets get a model going with two clusters and see what happens.

```{r Modeling Clusters}
# Create te model and visualize the results
km.cc <- kmeans(data, 2, nstart=25)
fviz_cluster(km.cc, data)
```

It's a little difficult to read, but I can see - generally speaking - what it's doing. We have our common core of spenders, and the outliers who are, for better or worse, more extravagant spenders. This can be seen by taking a look at the centroids for each variable. Let's do that really quick.

```{r Displaying Centroids}
# Display the cluster centers
km.cc$centers
```

Group 1 is our core group of spenders and you can see that they have lower balances, purchase amounts, purchase frequencies, etc. that the second group. Group 2 has markedly higher numbers for all those variables (they're spending more) while demonstrating lower percentages of full payments (they're more likely to carry a balance) over essentially the same tenure. One side note is that group 2 does have a significantly higher credit limit, which could be indicative of better spending habits, but could also be indicative of significantly higher income which does *not* necessarily imply better spending habits.
 
# FIN
 